Music Recommendation System Using the Million Song Dataset

### 2015-05-07 writeup/MSDgroupwriteup.tex
change in table
Update MSDgroupwriteup.tex

@@

\section{Conclusion}
After comparing the various algorithms, the simpler models gave us the best results. Our most complicated model, matrix factorization, gave us results barely above the popularity baseline. On the other hand, simply recommending users songs by artists they have already listened to (the artist-based baseline) resulted in a significantly better recommendation system. Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization. Our final results for each technique employed are summarized in the table below.

@@

\hline
\bf{Algorithm} &  \bf{MAP score}\\ \hline
Artist Based  Baseline  & 0.0448    \\ 
%\hline
Top 500 Songs by Count  Baseline &  0.0138  \\ 
%\hline
Top 500 Songs by Plays  Baseline &  0.0126  \\ 
%\hline
User Based  CF  &  0.0377 \\ 
%\hline
Item Based  CF  &  0.0479 \\ 
%\hline
Matrix Factorization  &   0.0143  \\ 
%\hline
\end{tabular}
\end{center}
\end{table}

>>>>

\citep{hu2008collaborative}

@@

\section{Results}
After comparing the various algorithms, the simpler models gave us the best results. Our most complicated model, matrix factorization, gave us results barely above the popularity baseline. On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system. Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization. Our final results for each technique employed are summarized in the table below.

@@

\hline
\bf{Algorithm} &  \bf{MAP score}\\ \hline
Artist Based Popularity Baseline  & 0.0448    \\ 
%\hline
Top 500 Songs by Count  Baseline &  0.0138  \\ 
%\hline
Top 500 Songs by Plays  Baseline &  0.0126  \\ 
%\hline
User Based  CF  &  0.0377 \\ 
%\hline
Item Based  CF  &  0.0479 \\ 
%\hline
Matrix Factorization  &   0.0143  \\ 
%\hline
\end{tabular}
\end{center}
\end{table}

\section{Conclusion}
Apart from the algorithms discussed above, we also explored building a recommendation system based on K-Means Clustering (by both user and song) and Non-Negative Matrix Factorization (NMF). In the end, both of these methods yielded worse results as compared to Probabalistic Matrix Factorization so we decided not to explore them further. Next steps include incorporating tags and metadata present in the dataset (e.g. year, genre, audio metadata). It would also be useful to expand the size of the subset and distribute the workload to multiple machines so that more data could be analyzed.
